Reliable feature point matching

### 2013-08-06 text/thesis.tex
Modified summary

\begin{figure}
	\centering%
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/MMC_pitts_source}
			\caption{Source image pair}
			\label{fig:pitts_source}
		\end{subfigure}%
		\\ %
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/MMC_pitts_keypoints}
			\caption{Feature points}
			\label{fig:pitts_keypoints}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		%\qquad (or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/mirror_match_off}
			\caption{\emph{SIFT}}
			\label{fig:unique}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/mirror_match_with_pruned}
			\caption{\emph{MM} intermediate result}
			\label{fig:within}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		%\qquad (or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/mirror_match}
			\caption{\emph{MM} final result}
			\label{fig:without}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/MMC_partition}
			\caption{\emph{MMC} Partition Example}
			\label{fig:pitts_partition}
		\end{subfigure}%
	\caption{Feature matching with \emph{MM} and \emph{MMC}. Dots represent feature points; green/red lines indicate correct/incorrect matches, respectively; black lines represent edges in the feature graph.  
	(c) Result of \emph{SIFT} matching. 
  (d) All matches found by \emph{MM}, including intra-image matches. 
	(e) Final \emph{MM} result. 
	(f) Example of a partition of feature points after clustering, which 
includes similar feature points from building windows and shirt patterns.}%
	\label{fig:comparemirror}%
\end{figure}%

@@

%			\caption{\emph{SIFT}}

@@

%(c) Result of \emph{SIFT} matching the two images.% 

@@

Figure~\ref{fig:comparemirror} illustrates the benefits of \emph{MM} 
using an example image pair from the Gallagher dataset 
\cite{gallagher2008}.
With \emph{SIFT} (Figure~\ref{fig:unique}), many incorrect matches occur 
in the fence towards the bottom of the image.
When we match all feature points together, many of these incorrect 
matches are eliminated, because points in the fence match with other 
points in the fence in the same image (Figures~\ref{fig:within} and
\ref{fig:without}).

@@

The matching algorithm for \emph{MMC}, \emph{getMatches}, finds matches 
within all partitions with more than two elements using the \emph{MM} 
approach.  However, as can be seen in the example in 
Figure~\ref{fig:cropped_graph}, many of the partitions contain only two 
feature points from different images linked by one edge. In such a case, 
we compare the similarity of the these two feature points with their 
second best match and remove matches where this ratio lies above a 
certain threshold, like in the \emph{SIFT} algorithm. For example in the 
case of Figure~\ref{fig:pitts_partition}, we have feature points from a 
building in one image grouped together with points from a shirt pattern 
in another.  The nearest neighbor method would have returned wrong 
matches, but since we match the partition with \emph{MM}, points in the 
building end up matching other points in the building, and no matches 
are returned.

@@

%
%
%\subsubsection{Experimental setup}
We compare the \emph{MM} and \emph{MMC} algorithms to 
\emph{SIFT} \cite{lowe2004sift} as well 
as \emph{Isodata} \cite{das2008event}. 
\emph{Isodata} uses geometric constraints, whereas \emph{SIFT} does not.
%, which uses geometric information to improve the matching. 
%The comparison  with \emph{SIFT} serves to illustrate the relative performance of 
%\emph{MM} and \emph{MMC} to a state of the art algorithm without 
%geometric constraints. \emph{Isodata} is included to 
%illustrate the performance of an algorithm relying on geometric 
%constraints faced with image pairs that might not have any overlap.  
The 
comparison was done using the \emph{Murals} dataset (Figure~\ref{fig:murals}), the 
\emph{Graf} set (Figure~\ref{fig:Graf}) from \cite{mikolajczyk2005performance}, and two 
images %\footnote{100\_1942.jpg and 100\_1941.jpg} 
from the Gallagher dataset \cite{gallagher2008} (Figure~\ref{fig:comparemirror}).

@@

Figure~\ref{fig:result_graf} shows the results for 100 patch pairs 
generated from images 1 and 3 of the \emph{Graf} image set (cf.~Figure~\ref{fig:Graf}). The plot 
compares the accuracy of \emph{SIFT}, \emph{MM}, \emph{MMC}, and 
\emph{Isodata} as a function of the number of matches produced (which is achieved by varying thresholds over a certain range). The results show that \emph{MM} and \emph{MMC} consistently outperform \emph{SIFT}; \emph{MMC} generally lies 2-3 percentage points above \emph{MM} when both are performing at optimal accuracy. 
Although \emph{Isodata} exhibits a good performance on strict thresholds (small number of matches), that quickly diminishes when more matches are desired.

@@

To further investigate the impact of viewpoint changes, we tested the 
algorithms on the \emph{Graf} image set (Figure~\ref{fig:Graf}), which 
contains 5 images of the same mural taken with gradually increasing 
viewpoint changes.  The first images are almost identical, while the 
last are taken from very different angles. The results from   \emph{MMC} 
and \emph{SIFT} as shown in Figure~\ref{fig:result_viewpoint}, 
confirming that \emph{MMC} is generally superior to \emph{SIFT} across 
viewpoint changes.
The performance of \emph{MM} (not shown in the plot) is similar to 
\emph{MMC}.

@@

In terms of computational complexity, \emph{SIFT}, \emph{MM}, and \emph{Isodata} can be 
implemented in $O(n\log n)$, where $n$ is the total number of feature 
points.  Our current \emph{MMC} implementation has a complexity of $O(n^2)$ due to the 
construction of a similarity matrix of the feature points. However, it is 
possible to approximate this in $O(n\log n)$ using search trees.

In terms of speed, Table~\ref{table:running_times} shows the 
running time of the four algorithms over 100 image pairs of $250\!\times\!250$ pixels. 
These numbers should be taken with a grain of salt, given that 
much of the code behind \emph{MMC} and \emph{Isodata} is implemented in 
Python, whereas \emph{MM} and \emph{SIFT} make use of OpenCV to execute 
computationally intensive operations in C++, which makes them much 
faster.

\begin{table}[htb]
\caption{Running times as tested on a Intel\textregistered\ Core\texttrademark\ i5-3550 CPU @ 
3.30~GHz with 8~GB memory.}
\label{table:running_times}
	\centering
%	\small
\begin{tabular}{r*{4}{c}}
\hline
	Algorithm: & \emph{SIFT} & \emph{MM} & \emph{MMC} & \emph{Isodata} 
	\\
	\noalign{\smallskip} 
	%
	Running time: & 20.94s & 18.23s & 883.99s & 555.33s \\
	\hline
\end{tabular}
\end{table}
%

@@

In many areas of computer vision matching local feature points is a key 
component, and matching algorithms have applications in problems such as
scene matching, object recognition, near duplicate detection and 
panorama recognition. To show the practical applications of the proposed 
matching algorithms, \emph{MM} and \emph{MMC}, we use show how they 
improve existing methods in near duplicate detection and panorama 
recognition. These two problems were chosen because they both involve 
matching images that might or might not overlap. These cases requires 
matching algorithms that limit false positives since e.g. returning the 
top $n$ matches for any given image pair would make it hard to 
distinguish adjacent or duplicate images from non-adjacent or dissimilar
image pairs. Near duplicate detection also constraints normal matching 
algorithms from relying on geometric assumptions, since we have very few
guarantees between two potential near duplicates that they should be 
matched using any particular geometric heuristic. In this section, we 
will first see \emph{MMC} applied to panorama recognition and alignment,
and next \emph{MMC} and \emph{MM} applied to near duplicate detection.

@@

\begin{figure}[h]
	\begin{subfigure}[t]{0.8\textwidth}
		\centering
		\includegraphics[width=\textwidth]{images/pano_narrow}
        \label{fig:pano_images}
		\caption{Source images}
    \end{subfigure}%
	\\
	\centering
	\begin{subfigure}{0.4\textwidth}
		\includegraphics[width=\textwidth]{images/panorama-autostitch}
        \label{fig:pano_autostitch}
		\caption{Panorama by Autostitch}
    \end{subfigure}%
	~%add desired spacing between images, e. g. ~, \quad, \qquad
	%(or a blank line to force the subfigure onto a new line)
	\begin{subfigure}{0.4\textwidth}
		\includegraphics[width=\textwidth]{images/panorama-MMC}
		\caption{Panorama by MMC}
		\label{fig:pano_MMC}
	\end{subfigure}%
    \label{fig:pano_example}
	\caption{An example of a difficult case for panorama recognition and 
		alignment.  In a) the source images used to generate the 
		panorama are displayed.\ b) shows the result of the Autostitch 
		program from \cite{brown2007automatic} which fails to combine 
		all images, while c) demonstrates the result of the simple 
		panorama stitcher
	using the proposed MMC algorithm. Here the two left images are 
correctly linked}
\end{figure}

@@

The results of the comparison can be seen in Table~\ref{table:nnd} where
\emph{MMC} improves performance by 8.05\% over the original framework 
using \emph{Ratio} to match images. For comparison, the performance 
improvement of \emph{PhotoCluster} compared to XMatch 
\cite{zhao2009scale}, the second best performing algorithm on the 
California-nd dataset according to \cite{vas2013cluster} is 2.52\%.

@@

We have addressed the problem of matching feature points without using 
geometrical constraints, proposing \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}.  The two algorithms 
share the common idea that feature points should have better 
matches in another image than in the image they came from to be 
considered good matches.  \emph{MMC} further improves on this idea by 
using the structure of the similarity graph of the feature points.

The algorithms show promising results when tested on the \emph{Murals} data 
set. %, where pairs of the same object seen from different angles are %cropped to have different degrees of overlap ranging from full to non at all. The results on 900 patch pairs show that 
\emph{MM} and \emph{MMC} 
generally outperform existing matching algorithms \emph{SIFT} and \emph{Isodata}, and \emph{MMC} 
outperforms \emph{MM}. We show that this result generalizes to 
variations in viewpoint change as well as more realistic photos featuring occlusions. 
% by comparing \emph{MMC} to \emph{SIFT} over a set of image pairs with an increasing magnitude of perspective difference. 
%We go on to apply the algorithms to a real life image case 
%featuring occlusion and a slight viewpoint change and show that the 
%performance of \emph{MM} and \emph{MMC} is consistent with the results 
%on \emph{Murals}.

Given the versatility of the proposed algorithms, we are planning to apply them to problems that require high 
reliability faced with images that might not match, such as near 
duplicate detection or face recognition.
%

>>>>

\begin{figure}
	\centering%
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/MMC_pitts_source}
			\caption{Source image pair}
			\label{fig:pitts_source}
		\end{subfigure}%
		\\ %
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/MMC_pitts_keypoints}
			\caption{Feature points}
			\label{fig:pitts_keypoints}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		%\qquad (or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/mirror_match_off}
			\caption{\emph{Ratio}}
			\label{fig:unique}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/mirror_match_with_pruned}
			\caption{\emph{MM} intermediate result}
			\label{fig:within}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		%\qquad (or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/mirror_match}
			\caption{\emph{MM} final result}
			\label{fig:without}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		\begin{subfigure}[t]{\columnwidth}
			\centering
			\includegraphics[width=0.85\columnwidth]{images/MMC_partition}
			\caption{\emph{MMC} Partition Example}
			\label{fig:pitts_partition}
		\end{subfigure}%
	\caption{Feature matching with \emph{MM} and \emph{MMC}. Dots represent feature points; green/red lines indicate correct/incorrect matches, respectively; black lines represent edges in the feature graph.  
	(c) Result of \emph{Ratio} matching. 
  (d) All matches found by \emph{MM}, including intra-image matches. 
	(e) Final \emph{MM} result. 
	(f) Example of a partition of feature points after clustering, which 
includes similar feature points from building windows and shirt patterns.}%
	\label{fig:comparemirror}%
\end{figure}%

@@

%			\caption{\emph{Ratio}}

@@

%(c) Result of \emph{Ratio} matching the two images.% 

@@

Figure~\ref{fig:comparemirror} illustrates the benefits of \emph{MM} 
using an example image pair from the Gallagher dataset 
\cite{gallagher2008}.
With \emph{Ratio} (Figure~\ref{fig:unique}), many incorrect matches occur 
in the fence towards the bottom of the image.
When we match all feature points together, many of these incorrect 
matches are eliminated, because points in the fence match with other 
points in the fence in the same image (Figures~\ref{fig:within} and
\ref{fig:without}).

@@

The matching algorithm for \emph{MMC}, \emph{getMatches}, finds matches 
within all partitions with more than two elements using the \emph{MM} 
approach.  However, as can be seen in the example in 
Figure~\ref{fig:cropped_graph}, many of the partitions contain only two 
feature points from different images linked by one edge. In such a case, 
we compare the similarity of the these two feature points with their 
second best match and remove matches where this ratio lies above a 
certain threshold, like in the \emph{Ratio} algorithm. For example in the 
case of Figure~\ref{fig:pitts_partition}, we have feature points from a 
building in one image grouped together with points from a shirt pattern 
in another.  The nearest neighbor method would have returned wrong 
matches, but since we match the partition with \emph{MM}, points in the 
building end up matching other points in the building, and no matches 
are returned.

@@

%
%
%\subsubsection{Experimental setup}
We compare the \emph{MM} and \emph{MMC} algorithms to 
\emph{Ratio} \cite{lowe2004sift} as well 
as \emph{Isodata} \cite{das2008event}. 
\emph{Isodata} uses geometric constraints, whereas \emph{Ratio} does not.
%, which uses geometric information to improve the matching. 
%The comparison  with \emph{Ratio} serves to illustrate the relative performance of 
%\emph{MM} and \emph{MMC} to a state of the art algorithm without 
%geometric constraints. \emph{Isodata} is included to 
%illustrate the performance of an algorithm relying on geometric 
%constraints faced with image pairs that might not have any overlap.  
The 
comparison was done using the \emph{Murals} dataset (Figure~\ref{fig:murals}), the 
\emph{Graf} set (Figure~\ref{fig:Graf}) from \cite{mikolajczyk2005performance}, and two 
images %\footnote{100\_1942.jpg and 100\_1941.jpg} 
from the Gallagher dataset \cite{gallagher2008} (Figure~\ref{fig:comparemirror}).

@@

Figure~\ref{fig:result_graf} shows the results for 100 patch pairs 
generated from images 1 and 3 of the \emph{Graf} image set (cf.~Figure~\ref{fig:Graf}). The plot 
compares the accuracy of \emph{Ratio}, \emph{MM}, \emph{MMC}, and 
\emph{Isodata} as a function of the number of matches produced (which is achieved by varying thresholds over a certain range). The results show that \emph{MM} and \emph{MMC} consistently outperform \emph{Ratio}; \emph{MMC} generally lies 2-3 percentage points above \emph{MM} when both are performing at optimal accuracy. 
Although \emph{Isodata} exhibits a good performance on strict thresholds (small number of matches), that quickly diminishes when more matches are desired.

@@

To further investigate the impact of viewpoint changes, we tested the 
algorithms on the \emph{Graf} image set (Figure~\ref{fig:Graf}), which 
contains 5 images of the same mural taken with gradually increasing 
viewpoint changes.  The first images are almost identical, while the 
last are taken from very different angles. The results from   \emph{MMC} 
and \emph{Ratio} as shown in Figure~\ref{fig:result_viewpoint}, 
confirming that \emph{MMC} is generally superior to \emph{Ratio} across 
viewpoint changes.
The performance of \emph{MM} (not shown in the plot) is similar to 
\emph{MMC}.

@@

In terms of computational complexity, \emph{Ratio}, \emph{MM}, and \emph{Isodata} can be 
implemented in $O(n\log n)$, where $n$ is the total number of feature 
points.  Our current \emph{MMC} implementation has a complexity of $O(n^2)$ due to the 
construction of a similarity matrix of the feature points. However, it is 
possible to approximate this in $O(n\log n)$ using search trees.

In terms of speed, Table~\ref{table:running_times} shows the 
running time of the four algorithms over 100 image pairs of $250\!\times\!250$ pixels. 
These numbers should be taken with a grain of salt, given that 
much of the code behind \emph{MMC} and \emph{Isodata} is implemented in 
Python, whereas \emph{MM} and \emph{Ratio} make use of OpenCV to execute 
computationally intensive operations in C++, which makes them much 
faster.

\begin{table}[htb]
\caption{Running times as tested on a Intel\textregistered\ Core\texttrademark\ i5-3550 CPU @ 
3.30~GHz with 8~GB memory.}
\label{table:running_times}
	\centering
%	\small
\begin{tabular}{r*{4}{c}}
\hline
	Algorithm: & \emph{Ratio} & \emph{MM} & \emph{MMC} & \emph{Isodata} 
	\\
	\noalign{\smallskip} 
	%
	Running time: & 20.94s & 18.23s & 883.99s & 555.33s \\
	\hline
\end{tabular}
\end{table}
%

@@

In many areas of computer vision matching local feature points is a key 
component, and matching algorithms have applications in problems such as
scene matching, object recognition, near duplicate detection and 
panorama recognition. To show the practical applications of the proposed 
matching algorithms, \emph{MM} and \emph{MMC}, we use show how they 
improve existing methods in near duplicate detection and panorama 
recognition. These two problems were chosen because they both involve 
matching images that might or might not overlap. These cases requires 
matching algorithms that limit false positives since e.g. returning the 
top $n$ matches for any given image pair would make it hard to 
distinguish adjacent or duplicate images from non-adjacent or dissimilar
image pairs. Near duplicate detection also constraints normal matching 
algorithms from relying on geometric assumptions, since we have very few
guarantees between two potential near duplicates that they should be 
matched using any particular geometric heuristic. In this section, we 
will first see \emph{MMC} applied to panorama recognition and alignment,
and next \emph{MM} and \emph{MMC} applied to near duplicate detection.

@@

\begin{figure}[h]
	\begin{subfigure}[t]{0.9\textwidth}
		\centering
        \includegraphics[width=\textwidth]{images/pano_wide}
        \label{fig:pano_images}
		\caption{Source images}
    \end{subfigure}%
	\\
	\centering
    \begin{subfigure}[b]{0.4\textwidth}
		\includegraphics[width=\textwidth]{images/panorama-autostitch}
        \label{fig:pano_autostitch}
		\caption{Panorama by Autostitch}
    \end{subfigure}%
	~%add desired spacing between images, e. g. ~, \quad, \qquad
	%(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.4\textwidth}
		\includegraphics[width=\textwidth]{images/panorama-MMC}
		\caption{Panorama by MMC}
		\label{fig:pano_MMC}
	\end{subfigure}%
    \label{fig:pano_example}
	\caption{An example of a difficult case for panorama recognition and 
		alignment.  In a) the source images used to generate the 
		panorama are displayed.\ b) shows the result of the Autostitch 
		program from \cite{brown2007automatic} which fails to combine 
		all images, while c) demonstrates the result of the simple 
		panorama stitcher
	using the proposed MMC algorithm. Here the two left images are 
correctly linked}
\end{figure}

@@

The results of the comparison can be seen in Table~\ref{table:nnd}. It 
shows that \emph{MMC} improves performance by 8.05\% over the original 
framework using \emph{Ratio} to match images. For comparison, the 
performance improvement of \emph{PhotoCluster} compared to XMatch 
\cite{zhao2009scale}, the second best performing algorithm on the 
California-nd dataset according to \cite{vas2013cluster}, is 2.52\%.

@@

This thesis has addressed the problem of matching feature points without 
using geometrical constraints, proposing \emph{Mirror Match (MM)} and 
\emph{Mirror Match with Clustering (MMC)}.  The two algorithms share the 
common idea that feature points should have better matches in another 
image than in the image they came from to be considered good matches.  
\emph{MMC} further improves on this idea by using the structure of the 
similarity graph of the feature points.

The algorithms show promising results when tested on the \emph{Murals} 
data set. The results on 900 image pairs show that \emph{MM} and 
\emph{MMC} generally outperform existing matching algorithms 
\emph{Ratio}, \emph{Isodata} and \emph{Spectral}. In addition \emph{MMC} 
outperforms \emph{MM}.  We show that this result generalizes to 
variations in viewpoint change by comparing \emph{MMC} to \emph{Ratio} 
over a set of image pairs with an increasing magnitude of perspective 
difference. We go on to apply the algorithms to a real life image case 
featuring occlusion and a slight viewpoint change and show that the 
performance of \emph{MM} and \emph{MMC} is consistent with the results 
on \emph{Murals}. To show that \emph{MM} or \emph{MMC}
can further improve the results of a geometrical method, we create 
\emph{Spectral-MMC} by combining \emph{Spectral} and \emph{MMC} and 
compare the three on the full \emph{Murals} dataset where it yields 
results better than both \emph{Spectral} and \emph{MMC}.

The practical application of \emph{MM} and \emph{MMC} is demonstrated by 
applying the algorithms to the problems of near duplicate detected and 
for the case of \emph{MMC} to panorama recognition and alignment. For 
the case of near duplicate detection, \emph{MMC} provides an 8.02\% 
improvement over the state of the art \emph{Ratio} in a near duplicate 
detection framework making use of local features. Similarly for panorama 
recognition and alignment a simple panorama stitching method based on 
\emph{MMC} succeeds in cases where Autostitch is unable to recognize and 
align the source images correctly. Both applications illustrate the 
central role that performance of the matching algorithm has on tasks 
that makes use of local features to match image pairs with potential 
overlap.
%
